### 2012-11-24 MSWL_ECO_lesson1.tex
Initial version latex files

\chapter{Introduction and motivation}

\section{Introduction}

\section{Part I}\label{Part I}  % TODO: Change for properly section title.

\section{Part II}\label{Part II} % TODO: Change for properly section title.

...

\section{Conclusions}\label{conclusions}